On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-26 TTau_Halpha/kurosawa.tex
Close to final version.
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@633 25464f21-6099-402c-b30a-640f1d40d38b

@@

  \caption{The effect of the mass-loss rate to mass-accretion rate
    ratio ($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$).  The
    H$\alpha$ profile computed with only the magnetospheric accretion
    (thick solid) is compared with the disc-wind-magnetosphere
    hybrid models for $\mu =$ 0.025 (dash-dot), 0.1
    (solid) and 0.2 (dash). The magnetosphere used here has
    $T_{\mathrm{max}}=7500\,\Kelvin$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$.  The
    temperature and the acceleration parameter of the disc wind are
    $T_{\mathrm{wind}}=9000\,\Kelvin$ and $\beta=0.5$ respectively.
    All the models are computed with the inclination angle
    $i=55^{\circ}$. As $\mu$ increases the emission component becomes
    stronger and the wings becomes broader, as expected. The
    absorption component is relatively insensitive to the value of
    $\mu$ in this example. The wide range of $\mu$ reasonably
    reproduces the profiles similar to ones seen in observations
    (e.g.~\citealt{reipurth:1996}).}

@@

@@

    \begin{tabular}{cc}
      \includegraphics[%
	clip,
	scale=0.45]{figures/discwind_acc_inc_effect_m9_b1_t7000.eps}&
      \includegraphics[%
	clip,
	scale=0.45]{figures/disc_wind_acc_inc_effect_m9_b2_t7000.eps}\tabularnewline
    \end{tabular}

@@

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the H$\alpha$ profiles computed using the alternative hybrid model at
the same inclination angles used for the disc-wind-magnetosphere
hybrid model. The wind acceleration parameter ($\beta$) and the wind
temperature ($T_{\mathrm{wind}}$)used here are $0.5$ and
$9000\,\Kelvin$ respectively. All the other parameters are same as in
the disc-wind magnetosphere model.  With this set of $\beta$ and
$T_{\mathrm{wind}}$, the wind emission is comparable or a slightly
greater than the emission from the magnetosphere. Because of the
geometry of the wind, the P-Cyg absorption feature weakens as the
inclination increases. The optical depth of the wind is much higher in
the polar direction compared to that at higher inclination. For the
same reason, the EW of the line increases as the inclination increase
as clearly seen in the figure.  Although not shown here, this tendency
of the inclination dependency holds for the models with this wind
models for wide ranges of $\beta$ and $T_{\mathrm{wind}}$ as long as
the emission from the wind is comparable to that of the magnetosphere.

@@

@@

The value of $\mu$, 0.1,  suggested by both MHD calculations and observations
(c.f.~\citep{koenigl:2000}) is consistent with the 
disc-wind-magnetosphere models presented here. We found that the
hybrid models with $0.025<\mu<0.2$ are also consistent with the type
of the profiles seen in the observations (e.g~\citealt{reipurth:1996}).

>>>>

To quantify the relative amount of the wind emission compared to the
that of the magnetosphere, we have computed the ratio the EW of the
hybrid model in Figure~\ref{fig:atlas_hybrid} with the EW of the
magnetosphere only model.   The EW ratio sometimes becomes below 1.0
when the wind only contributes to the absorption, but not to the
emission component of the resulting profile. 
In some models (e.g.~with $T_{\mathrm{wind}}=8000\,\Kelvin$ and
$\beta=2.0$), the contribution of the wind to the EW is much larger
than that from the magnetosphere (EW ratio is 3.2). This clearly
demonstrates that the difficulty of using the EW of H$\alpha$ as
accretion indicator since it might as the indicator for the
mass-loss.  We have also seen in Figure~\ref{fig:atlas_discwind_only}
that the disc wind models without the magnetosphere can produce the
H$\alpha$ line with significantly large EW. 
If the line fit was applied to an observed H$\alpha$ line by ignoring
the wind, the temperature and  
the mass-accretion rate of the magnetospherical accretion flow could 
be overestimated. A better approach to a model fit is a simultaneous
fitting of multiple lines. For example, the near-infrared lines such
as Pa$\beta$ and Br$\gamma$ should be used to determined the mass-loss
rate since these lines are much less affected by the wind. Using the
mass-loss rate from these lines, H$\alpha$ can be used to determine
the wind mass-loss rate.

@@

  \caption{The effect of the mass-loss rate to mass-accretion rate
    ratio ($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$).  The
    H$\alpha$ profile computed with only the magnetospheric accretion
    (thick solid) is compared with the disc-wind-magnetosphere
    hybrid models for $\mu =$ 0.05 (dash-dot), 0.1
    (solid) and 0.2 (dash). The magnetosphere used here has
    $T_{\mathrm{max}}=7500\,\Kelvin$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$.  The
    temperature and the acceleration parameter of the disc wind are
    $T_{\mathrm{wind}}=9000\,\Kelvin$ and $\beta=0.5$ respectively.
    All the models are computed with the inclination angle
    $i=55^{\circ}$. As $\mu$ increases the emission component becomes
    stronger and the wings becomes broader, as expected. The
    absorption component is relatively insensitive to the value of
    $\mu$ in this example. The wide range of $\mu$ reasonably
    reproduces the profiles similar to ones seen in observations
    (e.g.~\citealt{reipurth:1996}).}

@@

  \caption{The ratio of the line equivalent width (EW) of the
    disc-wind magnetosphere hybrid models in
    Figure~\ref{fig:atlas_hybrid} to the EW of the magnetosphere only model. }

\label{tab:ew_ratio}

\begin{tabular}{lccccc}
\hline 
$T_{\mathrm{max}}\,\,\left(\mathrm{K}\right)$&
&
$\beta$&
&
&
\tabularnewline
&
&
$0.2$&
$0.5$&
$1.0$&
$2.0$\tabularnewline
\hline 
$7000$&
&
$1.1$&
$.94$&
$.89$&
$.96$\tabularnewline
$8000$&
&
$1.1$&
$.92$&
$1.2$&
$3.2$\tabularnewline
$9000$&
&
$1.0$&
$1.1$&
$7.5$&
$36.$\tabularnewline
\hline
\end{tabular}

@@

    \begin{tabular}{cc}
      \includegraphics[%
	clip,
	scale=0.45]{figures/discwind_acc_inc_effect_m9_b1_t7000.eps}&
      \includegraphics[%
	clip,
	scale=0.45]{figures/stellarwind_inc_effect.eps}\tabularnewline
    \end{tabular}

@@

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the H$\alpha$ profiles computed using the alternative hybrid model at
the same inclination angles used for the disc-wind-magnetosphere
hybrid model. The wind acceleration parameter ($\beta$) and the wind
temperature ($T_{\mathrm{wind}}$)used here are $0.5$ and
$9000\,\Kelvin$ respectively. All the other parameters are same as in
the disc-wind magnetosphere model.  With this set of $\beta$ and
$T_{\mathrm{wind}}$, the wind emission is comparable or a slightly
greater than the emission from the magnetosphere. The profiles show
rather complex absorption features because of the interaction of
photons created in the wind with the magnetosphere. 
While the blue-shifted absorption features are formed in the wind, the
absorption features near the line centre and in the red wing (only for
$i=10^\circ$) are caused by the wind emission by the magnetosphere
which has a rather complex flow geometry. 
Because of the geometry of the wind, the P-Cyg absorption feature
weakens as the inclination increases. The optical depth of the wind is
much higher in 
the polar direction compared to that at higher inclination. For the
same reason, the EW of the line increases as the inclination increase
as clearly seen in the figure.  Although not shown here, this tendency
of the inclination dependency holds for the models with this wind
models for wide ranges of $\beta$ and $T_{\mathrm{wind}}$ as long as
the emission from the wind is comparable to that of the magnetosphere.

@@

The profiles computed with the hybrid model showed that the wind
emission contribution becomes significantly larger than that from the
magnetosphere (Figure~\ref{atlas_hybrid} and
Table~\ref{tab:ew_ratio}).  Henceforth, a large H$\alpha$ equivalent
width does not necessary indicate a large mass-accretion rate, but it
could be caused by the mass-loss by the wind outflow.

@@

The value of $\mu$, 0.1,  suggested by both MHD calculations and observations
(c.f.~\citep{koenigl:2000}) is consistent with the 
disc-wind-magnetosphere models presented here. We found that the
hybrid models with $0.05<\mu<0.2$ are also consistent with the type
of the profiles seen in the observations (e.g~\citealt{reipurth:1996}).
